---
title: "Supplement 5"
subtitle: "Risk factors"
author: "Peter Kamerman"
date: "Last updated: `r format(Sys.Date(), '%d %B %Y')`"
---

```{r setup, include = FALSE}
# Load packages
library(tidyverse)
library(magrittr)
library(DataExplorer)
library(rcompanion)
library(car)
library(sjPlot)
library(glmnet)
library(ResourceSelection)
library(patchwork)
library(rms)

# Define greyscale palette function (white to black)
grey_pal <- colorRampPalette(colors = c('#CCCCCC', '#000000'),
                             interpolate = 'linear')

pal <- grey_pal(3)

# Set ggplot theme
theme_set(new = theme_bw(base_size = 16))

# Knitr options
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = './figures/analysis-3/')
```

----

This script generates models for potential predictors for having pain.

Both univariate analyses and multi-variable model selection are presented.

----

# Import data

```{r import_data}
# Import data
pain <- read_rds('data-cleaned/wbpq.rds') %>% 
    select(PID,
           pain_in_last_month,
           pain_worst) %>% 
    mutate(pain = ifelse(pain_in_last_month == 'yes' & pain_worst > 0,
                         yes = 'yes',
                         no = 'no')) %>% 
    select(PID, pain)

general <- read_rds('data-cleaned/general_info.rds') %>% 
    select(PID, age, sex, educational_level, employment) %>% 
    mutate(employment = fct_collapse(employment,
                                     employed = c('employed', 'employed (part time)'),
                                     unemployed = c('unemployed'),
                                     grant = c('disability grant', 'pension grant')))

mental_health <- read_rds('data-cleaned/hscl.rds') %>% 
    select(PID, total_score)

# Join to core_info
data <- read_rds('data-cleaned/hiv_test.rds') %>%
    select(PID, test_result) %>% 
    left_join(pain) %>% 
    left_join(general) %>% 
    left_join(mental_health)
```

----

# Clean data

```{r clean}
# Remove participants without test results
data %<>%
    filter(!is.na(test_result))

# Remove participants with missing pain data
data %<>%
    filter(!is.na(pain))

# Convert character classes to factors
data %<>%
    mutate_if(is.character, factor)
```

----

# Quick look

```{r quick_look}
# Dataframe dimensions
dim(data)
# Column names
names(data)
# Glimpse data
glimpse(data)
```

----

# Check missingness

## Full cohort

```{r missingness}
data %>% 
    profile_missing() %>% 
    mutate(pct_missing = round(100 * pct_missing)) %>% 
    arrange(pct_missing)
```

## HIV-

```{r missingness2}
data %>% 
    filter(test_result == 'HIV negative') %>% 
    profile_missing() %>% 
    mutate(pct_missing = round(100 * pct_missing)) %>% 
    arrange(pct_missing)
```

## HIV+

```{r missingness3}
data %>% 
    filter(test_result == 'HIV positive') %>% 
    profile_missing() %>% 
    mutate(pct_missing = round(100 * pct_missing)) %>% 
    arrange(pct_missing)
```

----

# HIV status

## Build model
```{r hiv}
mod_hiv <- glm(pain ~ test_result,
               data = data,
               family = binomial(link = 'logit'))
```

## Beta coefficients
```{r hiv2}
# Coefficients
coef(mod_hiv)

# 95% CI of the coefficients
confint(mod_hiv)
```

## Odds ratio
```{r hiv3}
# OR
exp(coef(mod_hiv))

# 95% CI of the OR
exp(confint(mod_hiv))
```

## Overall model

```{r hiv4}
# likelihood ratio test
Anova(mod_hiv,
      test = 'LR')
```

## Model terms

```{r hiv5}
# Summary
summary(mod_hiv)

# Wald test
Anova(mod_hiv,
      type = 'II', 
      test = 'Wald')
```

## Model fit

### Pseudo-R^2

```{r hiv6}
nagelkerke(mod_hiv)
```

### Hosmer-Lemeshow test

```{r hiv7}
hoslem.test(x = mod_hiv$y,
            y = fitted(mod_hiv),
            g = 10)
```

## Plot predicted probabilities

```{r hiv8}
plot_model(mod_hiv, 
           type = 'pred')$test_result

# Publication plot
## Extract data
hiv <- plot_model(mod_hiv, 
                  type = 'pred')$test_result

hiv_data <- tibble(x = factor(hiv$data$x),
                   pred = hiv$data$predicted,
                   low = hiv$data$conf.low,
                   high = hiv$data$conf.high)

## Plot
pp_hiv <- ggplot(data = hiv_data) +
    aes(x = x,
        y = pred,
        ymin = low,
        ymax = high) +
    geom_errorbar(width = 0.3,
                  size = 1) +
    geom_point(size = 3) +
    annotate(geom = 'text',
             label = 'HIV status*',
             size = 5,
             x = 0.5, 
             y = 0.97,
             hjust = 0) +
    scale_y_continuous(limits = c(0, 1),
                       position = 'right') +
    scale_x_discrete(labels = c('Negative', 'Positive')) +
    labs(x = 'HIV test result') +
    theme(axis.title.y = element_blank(),
          axis.title.x = element_text(size = 17),
          panel.grid = element_blank(),
          axis.text = element_text(colour = '#000000'))
```

----

# Age

## Build model

```{r age}
mod_age <- glm(pain ~ age,
               data = data[!is.na(data$age), ],
               family = binomial(link = 'logit'))
```

## Beta coefficients

```{r age2}
# Coefficients
coef(mod_age)

# 95% CI of the coefficients
confint(mod_age)
```

## Odds ratios

```{r age3}
# OR
exp(coef(mod_age))

# 95% CI of the OR
exp(confint(mod_age))
```

## Overall model

```{r age4}
# Likelihood ratio test
Anova(mod_age,
      test = 'LR')
```

## Model terms

```{r age5}
# Summary
summary(mod_age)

# Wald test
Anova(mod_age,
      type = 'II', 
      test = 'Wald')
```

## Model fit

### Pseudo-R^2

```{r age6}
nagelkerke(mod_age)
```

### Hosmer-Lemeshow test

```{r age7}
hoslem.test(x = mod_age$y,
            y = fitted(mod_age),
            g = 10)
```

## Plot predicted probabilities
```{r age8}
plot_model(mod_age, 
           type = 'pred')$age 

# Publication plot
## Extract data
age <- plot_model(mod_age, 
                  type = 'pred')$age 

age_data <- tibble(x = age$data$x,
                   pred = age$data$predicted,
                   low = age$data$conf.low,
                   high = age$data$conf.high)

## Plot
pp_age <- ggplot(data = age_data) +
    aes(x = x,
        y = pred,
        ymax = high,
        ymin = low) +
    geom_ribbon(fill = '#CCCCCC') +
    geom_line(size = 0.8) +
    annotate(geom = 'text', 
             label = 'Age*', 
             size = 5,
             x = 10, 
             y = 0.97,
             hjust = 0) +
    scale_y_continuous(limits = c(0, 1),
                       position = 'left') +
    labs(x = 'Age (years)') +
    theme(axis.title.y = element_blank(),
          axis.title.x = element_text(size = 17),
          panel.grid = element_blank(),
          axis.text = element_text(colour = '#000000'))
```

----

# Sex

## Build model
```{r sex}
mod_sex <- glm(pain ~ sex,
               data = data[!is.na(data$sex), ],
               family = binomial(link = 'logit'))
```

## Beta coefficients

```{r sex2}
# Coefficients
coef(mod_sex)

# 95% CI of the coefficients
confint(mod_sex)
```

## Odds ratios

```{r sex3}
# OR
exp(coef(mod_sex))

# 95% CI of the OR
exp(confint(mod_sex))
```

## Overall model

```{r sex4}
# Likelihood ratio test
Anova(mod_sex,
      test = 'LR')
```

## Model terms

```{r sex5}
# Summary
summary(mod_sex)

# Wald test
Anova(mod_sex,
      type = 'II', 
      test = 'Wald')
```

## Model fit

### Pseudo-R^2
```{r sex6}
nagelkerke(mod_sex)
```

## Hosmer-Lemeshow test
```{r sex7}
hoslem.test(x = mod_sex$y,
            y = fitted(mod_sex),
            g = 10)
```

## Plot predicted probabilities
```{r sex8}
plot_model(mod_sex, 
           type = 'pred')

# Publication plot
## Extract data
sex <- plot_model(mod_sex, 
                  type = 'pred')$sex

sex_data <- tibble(x = factor(sex$data$x),
                   pred = sex$data$predicted,
                   low = sex$data$conf.low,
                   high = sex$data$conf.high)

## Plot
pp_sex <- ggplot(data = sex_data) +
    aes(x = x,
        y = pred,
        ymin = low,
        ymax = high) +
    geom_errorbar(width = 0.3,
                  size = 1) +
    geom_point(size = 3) +
    annotate(geom = 'text',
             label = 'Sex',
             size = 5,
             x = 0.5, 
             y = 0.97,
             hjust = 0) +
    scale_y_continuous(limits = c(0, 1),
                       position = 'right') +
    scale_x_discrete(labels = c('Female', 'Male')) +
    labs(x = 'Sex') +
    theme(axis.title.y = element_blank(),
          axis.title.x = element_text(size = 17),
          panel.grid = element_blank(),
          axis.text = element_text(colour = '#000000'))
```

----

# Educational level

## Build model

```{r school}
mod_school <- glm(pain ~ educational_level,
                 data = data[!is.na(data$educational_level), ],
                 family = binomial(link = 'logit'))
```

## Beta coefficients

```{r school2}
# Coefficients
coef(mod_school)

# 95% CI of the coefficients
confint(mod_school)
```

## Odds ratios

```{r school3}
# OR
exp(coef(mod_school))

# 95% CI of the OR
exp(confint(mod_school))
```

## Overall model

```{r school4}
# Likelihood ratio test
Anova(mod_school,
      test = 'LR')
```

## Model terms

```{r school5}
# Summary
summary(mod_school)

# Wald test
Anova(mod_school,
      type = 'II', 
      test = 'Wald')
```

## Model fit

### Pseudo-R^2

```{r school6}
nagelkerke(mod_school)
```

### Hosmer-Lemeshow test

```{r school7}
hoslem.test(x = mod_school$y,
            y = fitted(mod_school),
            g = 10)
```

## Plot predicted probabilities

```{r school8}
plot_model(mod_school, 
           type = 'pred')$educational_level +
    theme(axis.text.x = element_text(angle = 30, 
                                     hjust = 1))

# Publication plot
## Extract data
edu <- plot_model(mod_school, 
                  type = 'pred')$educational_level

edu_data <- tibble(x = factor(edu$data$x),
                   pred = edu$data$predicted,
                   low = edu$data$conf.low,
                   high = edu$data$conf.high)

## Plot
pp_edu <- ggplot(data = edu_data) +
    aes(x = x,
        y = pred,
        ymin = low,
        ymax = high) +
    geom_errorbar(width = 0.3,
                  size = 1) +
    geom_point(size = 3) +
    annotate(geom = 'text',
             label = 'Education*',
             size = 5,
             x = 0.5, 
             y = 0.97,
             hjust = 0) +
    scale_y_continuous(limits = c(0, 1),
                       position = 'left') +
    scale_x_discrete(labels = c('0-7', '8-12', '>12')) +
    labs(x = 'School grade') +
    theme(axis.title.y = element_blank(),
          axis.title.x = element_text(size = 17),
          panel.grid = element_blank(),
          axis.text = element_text(colour = '#000000'))
```

----

# Employment

## Build model

```{r employment}
mod_employment <- glm(pain ~ employment,
                      data = data[!is.na(data$employment), ],
                      family = binomial(link = 'logit'))
```

## Beta coefficients

```{r employment2}
# Coefficients
coef(mod_employment)

# 95% CI of the coefficients
confint(mod_employment)
```

## Odds ratios

```{r employment3}
# OR
exp(coef(mod_employment))

# 95% CI of the OR
exp(confint(mod_employment))
```

## Overall model

```{r employment4}
# Likelihood ratio test
Anova(mod_employment,
      test = 'LR')
```

## Model terms

```{r employment5}
# Summary
summary(mod_employment)

# Wald test
Anova(mod_employment,
      type = 'II', 
      test = 'Wald')
```

## Model fit

### Pseudo-R^2 

```{r employment6}
nagelkerke(mod_employment)
```

### Hosmer-Lemeshow test

```{r employment7}
hoslem.test(x = mod_employment$y,
            y = fitted(mod_employment),
            g = 10)
```

## Plot predicted probabilities

```{r employment8}
plot_model(mod_employment, 
           type = 'pred')$employment +
    theme(axis.text.x = element_text(angle = 30,
                                     hjust = 1))

# Publication plot
## Extract data
emp <- plot_model(mod_employment, 
                  type = 'pred')$employment

emp_data <- tibble(x = factor(emp$data$x),
                   pred = emp$data$predicted,
                   low = emp$data$conf.low,
                   high = emp$data$conf.high)

## Plot
pp_emp <- ggplot(data = emp_data) +
    aes(x = x,
        y = pred,
        ymin = low,
        ymax = high) +
    geom_errorbar(width = 0.3,
                  size = 1) +
    geom_point(size = 3) +
    annotate(geom = 'text',
             label = 'Employment',
             size = 5,
             x = 0.5, 
             y = 0.97,
             hjust = 0) +
    scale_y_continuous(limits = c(0, 1),
                       position = 'right') +
    scale_x_discrete(labels = c('Grant', 'Employed', 'Unemployed')) +
    labs(x = 'Employment status') +
    theme(axis.title.y = element_blank(),
          axis.title.x = element_text(size = 17),
          panel.grid = element_blank(),
          axis.text = element_text(colour = '#000000'))
```

----

# HSCL25 (total score)

## Build model

```{r hscl}
mod_hscl <- glm(pain ~ total_score,
                data = data[!is.na(data$total_score), ],
                family = binomial(link = 'logit'))
```

## Beta coefficients

```{r hscl2}
# Coefficients
coef(mod_hscl)

# 95% CI of the coefficients
confint(mod_hscl)
```

## Odds ratios

```{r hscl3}
# Odds ratio
exp(coef(mod_hscl))

# 95% CI of the OR
exp(confint(mod_hscl))
```

## Overall model

```{r hscl4}
# Likelihood ratio test
Anova(mod_hscl,
      test = 'LR')
```

## Model terms

```{r hscl5}
# Summary
summary(mod_hscl)

# Wald test
Anova(mod_hscl,
      type = 'II', 
      test = 'Wald')
```

## Model fit

### Pseudo-R^2 

```{r hscl6}
nagelkerke(mod_hscl)
```

### Hosmer-Lemeshow test

```{r hscl7}
hoslem.test(x = mod_hscl$y,
            y = fitted(mod_hscl),
            g = 10)
```

## Plot predicted probabilities

```{r hscl8}
plot_model(mod_hscl, 
           type = 'pred')

# Publication plot
## Extract data
hscl <- plot_model(mod_hscl, 
                   type = 'pred')$total_score 

hscl_data <- tibble(x = hscl$data$x,
                   pred = hscl$data$predicted,
                   low = hscl$data$conf.low,
                   high = hscl$data$conf.high)

## Plot
pp_hscl <- ggplot(data = hscl_data) +
    aes(x = x,
        y = pred,
        ymax = high,
        ymin = low) +
    geom_ribbon(fill = '#CCCCCC') +
    geom_line(size = 0.8) +
    annotate(geom = 'text', 
             label = 'HSCL-25', 
             size = 5,
             x = 1, 
             y = 0.97,
             hjust = 0) +
    scale_y_continuous(limits = c(0, 1),
                       position = 'left') +
    labs(x = 'HSCL-25 total score') +
    theme(axis.title.y = element_blank(),
          axis.title.x = element_text(size = 17),
          panel.grid = element_blank(),
          axis.text = element_text(colour = '#000000'))
```

```{r hscl_old, echo = FALSE, eval = FALSE}
# Publication plot
## Prepare data by binning frequency of pain into the nearest integer of 
## HSCL total score (e.g., 1.0 to 1.4 = 1, 1.5 to 2.0 = 2)
hscl <- data %>%
    select(pain, total_score) %>% 
    filter(complete.cases(.)) %>% 
    mutate(total_score = round(total_score, 1)) %>% 
    group_by(total_score, pain) %>% 
    summarise(count = n()) %>% 
    group_by(total_score) %>% 
    mutate(total = sum(count),
           freq = 100 * (count / total)) %>% 
    filter(pain == 'yes') %>% 
    mutate(total_score2 = round(total_score))

## Plot
pp_old <- ggplot(data = hscl) +
    aes(x = factor(total_score2), 
        y = freq) +
    geom_boxplot(size = 1) +
    scale_y_continuous(limits = c(0, 100)) +
    labs(x = 'HSCL25 total score',
         y = 'Frequency of pain (%)') +
    theme_bw(base_size = 20) +
    theme(panel.grid = element_blank(),
          axis.text = element_text(colour = '#000000'))

ggsave(filename = 'figures/figure2.png',
       plot = pp_old,
       width = 7,
       height = 6)
```

----

# Variable selection

## Using backward selection

### Prepare data

```{r bw}
# Extract complete cases and refactor educational levels
complete <- data %>%
    filter(complete.cases(.)) %>% 
    mutate(educational_level = fct_recode(educational_level,
                                          `1` = 'no/primary school',
                                          `2` = 'secondary school',
                                          `3` = 'post-school qualification'))

```

### Generate full model

```{r bw2}
fit <- lrm(pain ~ age + sex + test_result + total_score + educational_level + employment,
           data = complete, 
           x = TRUE, y = TRUE)
```

### Inspect full model coefficients 

```{r bw3}
# Model summary
fit

# Betas
coef(fit)
confint.default(fit)

# OR
exp(coef(fit))
exp(confint.default(fit))
```

### Perform backward selection on full model

```{r bw4}
# Perform selection
(bw  <- fastbw(fit))

# Betas
coef(bw)
confint.default(bw)

# OR
exp(coef(bw))
exp(confint.default(bw))
```

### Check model stability

100 bootstrapped resamples.

```{r bw5}
validate(fit, B = 100, bw = TRUE)
```

## Using LASSO

LASSO is a regression method that performs both variable selection and regularization in order to enhance the prediction accuracy and interpretability of the statistical model it produces.

The process involves performing a 10-fold cross validation to find the optimal _lambda_ (penalization parameter). And then running the analysis and extracting the model based on the best lambda.

- _lambda.min_ is the value of lambda that gives minimum mean cross-validated error.   

- _lambda.1se_, is the value of lambda that gives the most regularized model such that error is within one standard error of the minimum

### Generate a model matrix

```{r lasso}
# Extract complete cases
complete2 <- data %>% 
    filter(complete.cases(.))

# Dependent variable
y <- ifelse(complete2$pain == 'yes', yes = 1, no = 0) 

# Predictor variables
## Factor variables
xfactor <- model.matrix(complete2$pain ~ complete2$test_result + complete2$sex + 
                            complete2$educational_level + complete2$employment)[, -1]

## Combine with continuous variables
x <- as.matrix(data.frame(complete2$age, complete2$total_score, xfactor))
```

### Find the best minimum and 1SE lambda value using cross-validation

```{r lasso2}
# Set seed
set.seed(2019)

# Calculate lambda (alpha = 1, lasso)
cv.lasso <- cv.glmnet(x = x, y = y, 
                      nfolds = 10,
                      alpha = 1,
                      family = "binomial")

# Plot
plot(cv.lasso)
```

### Lambda values

#### Lambda min

```{r lasso3}
cv.lasso$lambda.min
```

#### Lambda 1se

```{r lasso4}
cv.lasso$lambda.1se
```

### Inspect the model coefficients

#### Lambda min

```{r lasso5}
# Betas
coef(cv.lasso, s = "lambda.min")

# OR
exp(coef(cv.lasso, s = "lambda.min"))
```

#### Lambda 1se

```{r lasso6}
# Betas
coef(cv.lasso, s = "lambda.1se")

# OR
exp(coef(cv.lasso, s = "lambda.1se"))
```

----

# Publication plot

```{r pub_plot}
composite_plot <- pp_hscl + pp_sex + pp_age + pp_hiv + pp_edu + pp_emp + 
    plot_layout(ncol = 2, nrow = 3)

ggsave(filename = 'figures/supp-figure-original.png',
       plot = composite_plot,
       height = 10,
       width = 8)
```

----

# Session information

```{r session_info}
sessionInfo()
```
